---
title: "Simpsons"
author: "Richard Luu"
date: '2022-09-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(tidyverse)
```


```{r setup.a, include=FALSE}
simps <- read_csv("data/simpsons_episodes.csv")
```

## Exploratory Data Analysis

### Observations

```{r EDA}
names(simps)
summary(simps)
```

- Most variables are in some way viable as predictor variable (i.e. description based off word/character appearance frequency)
- No variable for episode run time
- Does air date include time slot? (Relevance being primetime, Monday death slot for low rating shows etc.)
- Rating will be most interesting to analyze
- Refer to README for other preliminary questions
- Initial observations off of summary:
  - 724 episodes, 33 seasons
  - 1st air date, 12-17-1989 (If memory serves, it's Christmas special with Santa's Little Helper (family dog))
  - Error in # in series (Max 712713) and number episode in season (Max 1920?)
  - Lowest rating 4.0, but note IMDB ratings skewed / bias by nature
  
- Misc:

- Potential filter data by:
  - Tree house of Horror
  - First and last episode of season
- New external data points to consider:
  - Disney acquisition of Hulu / Fox, other sociopolitical/culture events (Presidential election episodes)

### Graphs

```{r Plot1}
ggplot(data = simps, aes(x = id, y = imdb_rating, col = as.factor(season))) + 
  geom_point() + ylab("Rating") + xlab("Episode Number") + ggtitle("Rating vs Episode")

temp1 <- ggplot(data = simps, aes(x = id, y = us_viewers_in_millions, col = as.factor(season))) + 
  geom_point() + ylab("Rating") + xlab("Episode Number") + ggtitle("Num Viewers vs Episode")


```

- General negative trend, however every season has a couple of outliers on both maxima
- Rating seems to almost plateau around mid 6
  - Start / end of season no notable patterns of yet 
  - i.e. season premiere vs finale (Presumably because of episodic nature and no overarching plotline)
- Seasons re-factored to better see trend as opposed to default gradient

- Changing predictor variable is interesting as pattern does not match way I thought it would
  - i.e. Lower ratings = lower viewer, but need to reflect about how sample size would affect ratings
  - Viewer count is in _millions_ 
  - Maybe an on/off season schedule? (Spring/Summer vs Fall/Winter)

```{r THOH}
#special_eps < simps %>% filter(simps, number_in_series = [4, 50, 20, 203])
THOH <- simps %>% filter(str_detect(title, 'Treehouse of Horror'))
ggplot(data = simps, aes(x = id, y = imdb_rating)) + geom_point() +
  geom_point(data = THOH, aes(x = id, y = imdb_rating, col = "")) + 
  ggtitle("Treehouse of Horror Specials")
```
- In general Treehouse of Horror (THOH) episodes seem to be on higher end of rating per season
- THOH 31 is lowest, and lower even within the immediate range / season

```{r Seasons}
get_season <- function(season_num) {
  tempDF <- simps %>% filter(season == season_num)
  return(tempDF)
}

s1 <- get_season(1)
head(s1)
```

## Seasonal Data EDA 


## Data Analysis

## Predictive Model

## Sentiment Analysis off Descriptions
