---
title: "Simpsons"
author: "Richard Luu"
date: '2022-09-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(tidyverse)
```


```{r setup.a, include=FALSE}
simps <- read_csv("C:/Users/richa/Desktop/Datasets/simpsons/simpsons_episodes.csv")
```

## Exploratory Data Analysis

### Observations

```{r EDA}
names(simps)
summary(simps)
```

- Most variables are in some way viable as predictor variable (i.e. description based off word/character apperance frequency)
- No variable for episode run time
- Does air date include time slot? (Relevance being primetime, Monday death slut for low rating shows etc.)
- Rating will be most interesting to analyze
- Refer to README for other preliminary questions
- Inital observations off of summary:
  - 724 episodes, 33 seasons
  - 1st air date, 12-17-1989 (If memory serves, it's Christmas special with Santa's Little Helper (family dog))
  - Error in # in series (Max 712713) and number episode in season (Max 1920?)
  - Lowest rating 4.0, but note IMDB ratings skewed / bias by nature

### Graphs

```{r Plot1}
ggplot(data = simps, aes(x = id, y = imdb_rating, col = as.factor(season))) + 
  geom_point() + xlab("Rating") + ylab("Episode Number") + ggtitle("Rating vs Episode")
```

- General negative trend, however every season has a couple of outliers on both maxima
- Rating seems to almost plateau around mid 6
  - Start / end of season no noteable patterns of yet 
  - i.e. season primere vs finale (Presumbily because of episodic nature and no overarching plotline)
- Seasons refactored to better see trend as opposed to default gradient


```{r Plot2}
#special_eps < simps %>% filter(simps, episodes = [])
```

- Note of community conisdered specials [Links](https://simpsonswiki.com/wiki/List_of_specials)
  - Scrape data
- New external data points to consider:
  - Disney acquistiono of Hulu / Fox, other sociopolitical/culture events (Presidential election episodes)
  - Special (voice) guest appearances (celebrities voicing themselves)