---
title: "simpsonsons Episodes Analysis"
author: "Richard Luu"
date: '2022-09-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(lattice)
library(grid)
library(glue) #
library(gridExtra)
library(tidyverse)
library(dplyr)
library(tibble)
library(stringr) #strspit
```

```{css, echo=FALSE}
<style type="text/css">
    ol { list-style-type: upper-alpha; }
</style>
```

```{r setup.a, include=FALSE}
simpsons <- read_csv("data/simpsons_episodes.csv")
```

# Data Source

- Data obtained from [Kaggle](https://www.kaggle.com/datasets/jonbown/simpsons-episodes-2016), but author does not mention where came from / how obtained
- Upon research, it's a mix of both Wikipedia and from IMDb (Internet Movie Database)
  - Prime example being viewership and production code from Wikipedia
  - While imdb_ratings explicit from IMDb (for obvious reasons)
- While investigating, Wikipedia also has Nielsen ratings
  - Can scrape and do t-tests to compare whether or not Nielsen ratings significantly different from IMDb


# Exploratory Data Analysis

## Observations

```{r EDA}
dim(simpsons)
names(simpsons)
summary(simpsons)
sum(is.na(simpsons))
head(as_tibble(simpsons))
```

- Most variables are in some way viable as predictor variable 
  - i.e. Use description text for text analysis (frequency, sentiment, etc.) as opposed to completely ignoring
- No variable for episode run time (Can assume ~22 mins per avg runtime for 30 minute tv slot)
- ~~Does air date include time slot? (Relevance being primetime, [Friday death slot](https://en.wikipedia.org/wiki/Friday_night_death_slot) for low rating shows pending cancellation)~~
  - Upon further inspection, not relevant 
- Rating will be most interesting to analyze/use as independent variable
- Variables changed for ease of access
  - `imdb_rating` to `rating`
  - `us_viewers_in_millions` to `viewers`

```{r New-Var}
desc_split <- strsplit(simpsons$description, " ")
length_description <- sapply(desc_split, function(x) sum(str_count(x, "\\S+"))) 
age <- difftime("2022-09-30", simpsons$original_air_date)
simpsons$length_description <- length_description
simpsons$age <- age
```

```{r Length_desc}
summary(simpsons$length_description)
```
  
- Adding two new variables:
  - `length_description` and `age` (Time since original air date)
    - At time of starting this (Late September), Season 34 already airing so if any episode `age` anylsis is conducted will be based off this date as newer seasons [start around this time](https://en.wikipedia.org/wiki/List_of_The_Simpsons_episodes) 
    - Shortest episode description is 5 words:
      - `which(simpons$length_descrtipion == 5)`
      - 3 cases (187, 295, 714)

```{r data-rename, include=FALSE}
simpsons <- simpsons %>% dplyr::rename(viewers = us_viewers_in_millions)
simpsons <- simpsons %>% rename(rating = imdb_rating)
simpsons <- simpsons %>% rename(directors = directed_by)
simpsons <- simpsons %>% rename(writers = written_by)
```

- Initial observations off of summary:
  - 725 x 12 table
  - 724 episodes, 33 seasons
  - 1st air date, 12-17-1989 
    - If memory serves, it's Christmas special with Santa's Little Helper (family dog)
  - Lowest rating 4.0, but note IMDB ratings skewed / bias by nature
  
Potential errors in data based off summary:
1. Error in `number_in_series` (Max 712713)
2. Number episode in season (Max 1920) 
3. No blatant missing values but some numeric values listed as different type 
  - (us_viewers_in_million) is `char` type

* Refer to README for other preliminary questions
  
- Misc:
  - Potential filter data by:
    - Treehouse of Horror
    - First and last episode of season
  - New external data points to consider:
    - Disney acquisition of Hulu / Fox and FX
    - Sociopolitical/culture events (i.e. Presidental election episodes)

## Data Error Discussion

### A. Number In Series
```{r asd}
head(arrange(simpsons, desc(number_in_series)))
subsetDF <- c("id", "title", "production_code", "directors", "writers", "number_in_season", "number_in_series")
simpsons_sub <- simpsons[,subsetDF]
head(simpsons_sub)
```
- Sorted by max `number_in_series` in which reveals 3 cases where records were wrong
  - Number in series: 607, 679, 709
  - Other columns also stood out, and subsetted accordingly for further investigation

### B. Number in Season
```{r Misc Data}
# simpsons[which.max(simpsons$number_in_season),]

# head(arrange(simpsons, desc(number_in_season)))
```
- Two errors where the episode number in season is over 1k 
- Interesting to note that both eps are also wrong in num in series
  - Assuming the data is scrapped off of Wikipedia, upon further inspection they are 2-part episode(s)
    - Scrapped to where information is concatenated together
      - i.e.`number_in_season` is `1213`, it's a 2 parter of episodes 12 and 13 respectively
      - Note: Episode 13 does not _officially_ exist
      - If need to reference by `number_in_series` use `id`, however `number_in_season` fixed for season based analysis
    - Noted in episode titles, multiple writers, and different production codes

```{r Data-Fix}
simpsons[607,]$number_in_season = 12
simpsons[679,]$number_in_season = 19
simpsons[709,]$number_in_season = 6
```
  
### C. Rating
- Simply refactor by `as.numeric()` for further application

- - - 

## Graphs

```{r BoxPlots-Seasons}
(ggplot(data = simpsons, aes(x = as.factor(season), y = rating, color=as.factor(season))) + geom_boxplot() + ggtitle("Ratings")) 

(ggplot(data = simpsons, aes(x = as.factor(season), y = as.numeric(viewers), color=as.factor(season))) + geom_boxplot() + ggtitle("Viewers"))
```

Note: Graphs expected to fall under time series analysis, and using id/original_air_date/age are interchangeable since they are effectively factors/'categorical.'


```{r Rating vs Episode}
(tempPlot <- ggplot(data = simpsons, aes(x = id, y = rating, col = as.factor(season), shape=as.factor(season))) + 
  geom_point() + ylab("Rating") + xlab("Episode Number") + ggtitle("Rating vs Episode") + 
  geom_line() + geom_line() + 
  scale_shape_manual(values = rep(0:14, 3))) 

ggplot(data = simpsons, aes(x=id, y = rating)) + geom_point()
ggplot(data = simpsons, aes(x=original_air_date, y = rating)) + geom_point()
```

- General negative trend, however every season has a couple of outliers on both maxima
- Rating seems to almost plateau around mid 6
  - Start / end of season no notable patterns of yet 
  - i.e. season premiere vs finale (Presumably because of episodic nature and no overarching plotline)
- Seasons re-factored to better see trend as opposed to default gradient


```{r (old) Viewers vs Episode}
(oldViewers <- ggplot(data = simpsons, aes(x = id, y = viewers, col = as.factor(season))) + 
  geom_point() + ylab("Viewers") + xlab("Episode Number") + ggtitle("Num Viewers vs Episode"))
```

- ~~~Changing predictor variable to viewer count is interesting as pattern does not match way I thought it would~~~
  - ~~~i.e. Lower ratings = lower viewer, but need to reflect about how sample size would affect ratings~~~
  - Viewer count is in _millions_ 
  - ~~~Maybe an on/off season schedule? (Spring/Summer vs Fall/Winter)~~~
  - NOTE: Some data is missing are some `us_viewers_in_millions` values hard-coded as `N/A` (char class)
    - 159, 160, 172

```{r Viewers Fixed}
simpsons$viewers <- as.numeric(simpsons$viewers)
(refactoredViewer <- ggplot(data = simpsons, aes(x = id, y = viewers, col = as.factor(season))) + 
  geom_point() + ylab("Viewers") + xlab("Episode Number") + ggtitle("Num Viewers vs Episode") +
  scale_y_continuous(name="Viewers (in mil)", limits=c(0,30)))
```

- Upon refactoring viewer count to numeric trend does indeed follow expectations

```{r THOH}
#special_eps < simpsons %>% filter(simpsons, number_in_series = [4, 50, 20, 203])
THOH <- simpsons %>% filter(str_detect(title, 'Treehouse of Horror'))
ggplot(data = simpsons, aes(x = id, y = rating)) + geom_point() +
  geom_point(data = THOH, aes(x = id, y = rating, col = "")) + 
  ggtitle("Treehouse of Horror Specials")

ggplot(data = simpsons, aes(x = id, y = as.numeric(viewers))) + geom_point() +
  geom_point(data = THOH, aes(x = id, y = as.numeric(viewers), col = "")) + 
  ggtitle("Treehouse of Horror Specials")
```
- In general Treehouse of Horror (THOH) episodes seem to be on higher end of rating per season
- THOH 31 is lowest, and lower even within the immediate range / season

### Seasonal Data EDA  (+ Season length span)

```{r Seasons}
get_season <- function(season_num) {
  tempDF <- simpsons %>% filter(season == season_num)
  return(tempDF)
}

set1 <- c()
set2 <- c()
set3 <- c()
# 
# plot_data_column = function (data) {
#     ggplot(data, aes_string(x = number_in_season, y = as.numeric(us_viewers_in_millions))) +
#         geom_point()
# }
# 
# myplots <- lapply(1:12, plot_data_column, data = get_season(i))

for (i in 1:12) {
  temp <- get_season(i)
  tempGG <- ggplot(temp, aes(x = id, y = viewers)) + geom_point() + 
    ggtitle(glue('Season {i}'))
  set1[[i]] = tempGG
}

tempIndex = 1
for (i in 13:24) {
  temp <- get_season(i)
  tempGG <- ggplot(temp, aes(x = id, y = viewers)) + geom_point() + 
    ggtitle(glue('Season {i}'))
  set2[[tempIndex]] = tempGG
  tempIndex = tempIndex + 1
}

tempIndex = 1
for (i in 24:max(simpsons$season)) {  
  temp <- get_season(i)
  tempGG <- ggplot(temp, aes(x = id, y = viewers)) + geom_point() + 
    ggtitle(glue('Season {i}'))
  set3[[tempIndex]] = tempGG
  tempIndex = tempIndex + 1
}

# grid.arrange(set1[[1]], set1[[2]], set1[[3]], set1[[4]], ncol=2, nrow=2)

grid.arrange(grobs = set1, ncol = 3, nrow=4)
grid.arrange(grobs = set2, ncol = 3, nrow=4)
grid.arrange(grobs = set3, ncol = 3, nrow=4)
# 
# for (i in 1:10) {
#   print(set1[i])
# }
# s1 <- get_season(1)
# ggplot(s1, aes(x = id, y = as.numeric(us_viewers_in_millions))) + geom_point()
# head(s1)
```



## Writers / Directors

```{r Writers}
length(unique(simpsons$writers))
length(unique(simpsons$directors))


ggplot(simpsons, aes(writers, fill = writers)) + geom_bar() +
  coord_polar("y", start = 0)
```

- 187 unique writers records (Some entries have multiple people, i.e. outliers mentioned above - 2 part episodes)
- 57 unique director listings 

## Data Analysis

## Predictive Model

## Sentiment Analysis off Descriptions



#----------------------

```{r}
tempSeason1 <- get_season(1)
tempSeason2 <- get_season(2)


ggplot(data=simpsons, aes(x = original_air_date, y = rating)) + geom_point(aes(col = as.factor(season))) + geom_smooth(data = simpsons, method=lm, se=FALSE, fullrange=FALSE, aes(group=as.factor(season), col="#000000"))
  
  geom_smooth(data = get_season(1), aes(x=original_air_date, y = imdb_rating)) 

+ 
  geom_smooth(data = get_season(2), aes(x=original_air_date, y = imdb_rating, col="2")) 




```


# Misc Resources
- [strsplit()](https://stackoverflow.com/questions/8920145/count-the-number-of-all-words-in-a-string)
- [Collapsable Elements](https://stackoverflow.com/questions/52576626/rmarkdown-collapsible-panel)
