---
title: "Simpsons Episodes Analysis"
author: "Richard Luu"
output: 
  html_document:
    toc: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(lattice)
library(grid)
library(glue) #
library(gridExtra)
library(tidyverse)
library(dplyr)
library(tibble)
library(mice) #md.pattern
library(stringr) #strspit
library(plotly)
library(DT)
```

```{css, echo=FALSE}
<style type="text/css">
    ol { list-style-type: upper-alpha; }
</style>
```

```{r data_load, include=FALSE}
simpsons <- read_csv("data/simpsons_episodes.csv")
```

# Preface
- Specific references to season/episode short hand is: S#E# (Season #, Episode #)
- Most plots interactive / have hover-able elements to inspect specific values 
- Observational claims not statistically proven until said otherwise
  - i.e. "The average rating of season 2 is greater than season 7" While yes the plots show it but it is not statistically accurate / true

# Data Source

- Data obtained from [Kaggle](https://www.kaggle.com/datasets/jonbown/simpsons-episodes-2016), but author does not mention where came from / how obtained
- Upon research, it's a mix of both Wikipedia and IMDb (Internet Movie Database)
  - Prime example being viewership and production code from Wikipedia
  - While `imdb_ratings` explicit from IMDb (for obvious reasons)
- While investigating, Wikipedia also has Nielsen ratings
  - Can scrape and do t-tests to compare whether or not Nielsen ratings significantly different from IMDb

# Exploratory Data Analysis

## Observations

```{r class.source = 'fold-show', EDA.1}
dim(simpsons)
names(simpsons)
summary(simpsons)
sum(is.na(simpsons))
```

```{r EDA.2}
head(simpsons) %>% rmarkdown::paged_table()
```

```{r new_vars, include=FALSE}
desc_split <- strsplit(simpsons$description, " ")
length_description <- sapply(desc_split, function(x) sum(str_count(x, "\\S+"))) 
age <- difftime("2022-09-30", simpsons$original_air_date)
simpsons$length_description <- length_description
simpsons$age <- age
```

```{r data_rename, include=FALSE}
#simpsons <- simpsons %>% dplyr::rename(viewers = us_viewers_in_millions)
simpsons <- simpsons %>% rename(rating = imdb_rating)
simpsons <- simpsons %>% rename(directors = directed_by)
simpsons <- simpsons %>% rename(writers = written_by)
simpsons$season <- as.factor(simpsons$season)
```

```{r Length_desc}
summary(simpsons$length_description)
```
  
- Most variables are in some way viable as predictor variable
  - i.e. Use description text for text analysis (frequency, sentiment, etc.) as opposed to completely ignoring
- No variable for episode run time (Can assume ~22 mins per avg run time for 30 minute tv slot)
- ~~Does air date include time slot? (Relevance being primetime, [Friday death slot](https://en.wikipedia.org/wiki/Friday_night_death_slot) for low rating shows pending cancellation)~~
  - Upon further inspection, shouldn't be _too_ relevant 
- Rating will be most interesting to analyze/use as independent variable
- Variables changed for ease of access
  - `imdb_rating` to `rating`
  - `us_viewers_in_millions` to `viewers`
  
  
- Adding two new variables:
  - `length_description` and `age` (Time since original air date)
    - At time of starting this analysis, Season 34 has already started airing so if any episode `age` analysis is conducted will be based off this date as newer seasons [start around this time](https://en.wikipedia.org/wiki/List_of_The_Simpsons_episodes) 
    - Shortest episode description is 5 words:
    ```{r}
    simpsons$description[which(simpsons$length_description == 5)]
    ```
      - 3 cases (187, 295, 714)

**Initial observations off of summary**:

  - 725 x 12 table
  - 724 episodes, 33 seasons
  - 1st air date, 12-17-1989 
    - If memory serves, it's Christmas special with Santa's Little Helper (family dog)
  - Lowest rating 4.0, but note IMDB ratings skewed / bias by nature
  
Potential errors in data based off summary:

1. Error in `number_in_series` (Max 712713)

2. Number episode in season (Max 1920) 

3. No blatant missing values but some numeric values listed as different type 
    - (`us_viewers_in_million`) is `char` type

Refer to README for other preliminary questions
  
- Misc:
  - Potential filter data by:
    - Treehouse of Horror
    - First and last episode of season
  - New external data points to consider:
    - Disney acquisition of Hulu / Fox and FX
      - Fox: March 19, 2019
      - "In March 2019, Disney acquired 21st Century Fox, giving it a 60% majority stake in Hulu"
      https://en.wikipedia.org/wiki/Hulu#Disney_majority_ownership
    - Sociopolitical/culture events (i.e. Presidental election episodes)
    - Simpsons movie release date: July 27, 2007

## Data Error Discussion

### A. Number In Series
```{r asd}
rmarkdown::paged_table(head(arrange(simpsons, desc(number_in_series))))

#DT::datatable(head(arrange(simpsons, desc(number_in_series))))
subsetDF <- c("id", "title", "production_code", "directors", "writers", "number_in_season", "number_in_series")
simpsons_sub <- simpsons[,subsetDF]
```

- Sorted by max `number_in_series` in which reveals there cases where records were wrong
  - Number in series/id: 607, 679, 709
  - While scrolling through the table, other columns also stood out (and subsetted by following variables accordingly for further investigation)
  
```{r subset_data}
subsetDF
```

### B. Number in Season
```{r Misc Data}
#simpsons[which.max(simpsons$number_in_season),]
#head(arrange(simpsons, desc(number_in_season)))
rmarkdown::paged_table(head(arrange(simpsons_sub[c("id", "title","number_in_season")], desc(number_in_season))))

```
- Two errors where the episode number in season is over 1k 
- Interesting to note that both eps are also wrong in `num_in_series`
  - Assuming the data is scrapped off of Wikipedia, upon further inspection they are 2-part episode(s) (also noted by episode title)
    - Scrapped to where information is concatenated together
      - i.e.`number_in_season` is `1213`
        - 2 part episode, of episodes 12 and 13 respectively
      - Note: Episode 13 does not _officially_ exist
      - If need to reference by `number_in_series` use `id`, however `number_in_season` adjusted accordingly as two episodes
    - This concatenation in episode titles, multiple writers, and different production codes data differs
- Since it's a two-part episode that premiered together, can assume same concurrent viewers for both part on same date
- Raises issue of should the episodes be treated separately as different parts, or one whole? For sake of consistency, they will be two separate entities (since they have different writers and production codes)

```{r Data-Fix, include=FALSE}
#simpsons[607,]$number_in_season = 12
#simpsons[679,]$number_in_season = 19
#simpsons[709,]$number_in_season = 6
old_simps <- simpsons
```
  
### C. Viewers
```{r class.source = 'fold-show', numeric-viewers}
simpsons$viewers <- as.numeric(simpsons$us_viewers_in_millions)
which(is.na(as.numeric(simpsons$viewers)))
simpsons[c(160, 161, 173),c("us_viewers_in_millions", "viewers")]
simpsons_rm <- simpsons[-c(160, 161, 173),]
```

- Simply refactor by `as.numeric()` for further application
- Missing data points, originally missed because of type issue
  - `is.na()` didn't detect because it was of class `char`
  - Data is missing from Wikipedia
    - Interesting observation made -- Wikipedia articles report viewership in both (presumed) individual viewers and viewing households
      - i.e. Compare the table of episode / viewer to individual episode rating under reception for Nielsen rating and viewing households
      - Ratio comparing existing data of viewers to viewing household came to be about (1.7~1.8):1 (for case of season 8)
  - Season: 8, Episode(s): 7,8,20
- [Multiple ways to treat missing data for forecasting](https://data.library.virginia.edu/getting-started-with-multiple-imputation-in-r/)
  - Will try all 3, will start out with deleting missing rows. 
  - Noted mean calculations earlier can be used for method 2.
  - Inputation will use other packages.


(Match id to 607, 679, 709)

```{r data-fix, include=FALSE}
tempSimps <- simpsons_rm
#simpsons[607,]$number_in_season = 12
#simpsons[679,]$number_in_season = 19
#simpsons[709,]$number_in_season = 6

#(temp607 <- simpsons[608,])
(phatsby_vol1 <- tempSimps[tempSimps$id == 607,])
phatsby_vol1$id = 607
phatsby_vol1$title = "The Great Phatsby (Vol. I)"
phatsby_vol1$production_code = "WABF04"
phatsby_vol1$directors = "Chris Clements"
phatsby_vol1$writers = "Dan Greaney"
phatsby_vol1$number_in_season = 12
phatsby_vol1$number_in_series = 608

phatsby_vol2 <- phatsby_vol1
phatsby_vol2$id = 607
phatsby_vol2$title = "The Great Phatsby (Vol. II)"
phatsby_vol2$production_code = "WABF05"
phatsby_vol2$directors = "Timothy Bailey"
phatsby_vol2$writers = "Matt Selman"
phatsby_vol2$number_in_season = 13
phatsby_vol2$number_in_series = 609

tempSimps[tempSimps$id == 607,] = phatsby_vol1
tempSimps[nrow(tempSimps) + 1, ] = phatsby_vol2

#----------------

(warrin_priests1 <- tempSimps[tempSimps$id == 679,])
warrin_priests1$id = 679
warrin_priests1$title = "Warrin' Priests (Part One)"
warrin_priests1$production_code = "ZABF12"
warrin_priests1$directors = "Bob Anderson"
warrin_priests1$writers = "Pete Holmes"
warrin_priests1$number_in_season = 19
warrin_priests1$number_in_series = 681

warrin_priests2 <- warrin_priests1
#warrin_priests2$id = 680
warrin_priests2$title = "Warrin' Priests (Part Two)"
warrin_priests2$production_code = "ZABF13"
warrin_priests2$directors = "Matthew Nastuk"
#Same writer
warrin_priests2$number_in_season = 20
warrin_priests2$number_in_series = 682

tempSimps[tempSimps$id == 679,] <- warrin_priests1
tempSimps[nrow(tempSimps) + 1, ] <- warrin_priests2

#----------------
(simpsons[710,])
srs_flanders1 <- tempSimps[tempSimps$id == 709,]
srs_flanders1$id = 709
srs_flanders1$title = "A Serious Flanders (Part One)"
srs_flanders1$production_code = "QABF21"
srs_flanders1$directors = "Debbie Bruce Mahan"
srs_flanders1$writers = "Cesar Mazariegos"
srs_flanders1$number_in_season = 6
srs_flanders1$number_in_series = 712

srs_flanders2 <- srs_flanders1
srs_flanders2$title = "Warrin' Priests (Part One)"
srs_flanders2$production_code = "QABF22"
srs_flanders2$directors = "Matthew Faughnan"
srs_flanders2$number_in_season = 7
srs_flanders2$number_in_series = 713

tempSimps[tempSimps$id == 709,] <- srs_flanders1
tempSimps[nrow(tempSimps) + 1, ] <- srs_flanders2
```

- - - 

## Graphs

Note: Graphs expected to fall under time series analysis, and using `id`/`original_air_date`/`age` are somewhat interchangeable since they are effectively factors/'categorical.'

### Seasonal 

#### Ratings

```{r BoxPlots-SeasonsVsRating}
rating_v_sason <- ggplot(data = tempSimps, aes(x = as.factor(season), y = rating, color = as.factor(season)))

temp <- ggplot(data = tempSimps, aes(x = as.factor(season), y = rating, color=as.factor(season))) + geom_boxplot() + ggtitle("IMDB Rating vs Simpsons Seasons") + xlab("Season") + ylab("Rating") + theme(legend.position = "none") +
  scale_y_continuous(name="IMDB Rating",limits=c(4, 10))

ggplotly(temp)

sznAvg <- tempSimps %>% group_by(season) %>% summarise(sznAvg = mean(rating))

temp2 <- ggplot(data = sznAvg, aes(x = season, y = sznAvg)) + geom_point() + ggtitle("Average IMDB Rating vs Simpsons Season") + xlab("Season") + ylab("Average IMDB Rating")

ggplotly(temp2)
```

- Outliers in this case are more interesting to look at when compared to `viewers`
- Ratings seem to plateau as seasons progress, but the occasional outlier does exist
  - Doesn't necessarily mimic viewership count
    - Note, viewership could also decrease with growth of streaming bubble -- people are not pressed to watch shows live / on premiere unless a devout fans / avid watcher
- Seasons with large spread: 8, 23, and potentially 32
- S8 has highest maximum, alterantively S23 lowest:
  - S8E23: [Homer's Enemy](https://en.wikipedia.org/wiki/Homer%27s_Enemy)
  - S23E23: [Lisa Goes Gaga](https://en.wikipedia.org/wiki/Lisa_Goes_Gaga)
  
- When looking solely at the averages, the declining pattern is more definite
- Season 30 has a large drop off from 29, then recovers the following seasons
  - comparable to season 8 to 9
  
```{r viewers-outliers, include=FALSE}
tempSimps[tempSimps$season == 16,]
tempSimps[tempSimps$id == 342,]
```

#### Viewers

```{r BoxPlot-SeasonVsViewer}
tempViewers <- ggplot(data = tempSimps, aes(x = as.factor(season), y = as.numeric(viewers), color=as.factor(season))) + geom_boxplot() + ggtitle("Viewers vs Simpsons Seasons") + xlab("Season") + ylab("Viewers (p/ mil)") +theme(legend.position = "none")

ggplotly(tempViewers)
```
- Viewership does seemingly appear to decrease over time
- More outliers exist for maxima rather than minima
  - Two cases (Season 1 and 13) where the outlier are minima
  - Most extreme case of maximum in [S16 E8](https://en.wikipedia.org/wiki/Homer_and_Ned%27s_Hail_Mary_Pass) (id = 342)
    - Interesting to see how rest of maxima not even close to same viewership
    - Gap b/w the maximum and q3 largest compared to rest of seasons
    - Why is it this specific episode that makes it an outlier?
      - Note of cultural references within Wikipedia, but the viewers would have no way of knowing prior unless teased
      - https://www.onthisday.com/date/2005/february/6
      - Same date of superbowl (also noted in header of Wikipedia page)
    
- https://en.wikipedia.org/wiki/Homer_and_Ned%27s_Hail_Mary_Pass

### Episodic Data

#### Ratings

```{r Rating vs Episode}
tempPlot <- ggplot(data = tempSimps, 
                   aes(x = id, y = rating, col = as.factor(season), shape=as.factor(season))) + geom_point() + 
  ylab("Rating") + xlab("Episode Number") + ggtitle("Rating vs Episode") + 
  geom_line() + geom_line() + scale_shape_manual(values = rep(0:14, 3)) +
   theme(legend.position = "none") + scale_y_continuous(name="IMDB Rating", limits=c(4,10))

ggplotly(tempPlot)

#ggplot(data = simpsons, aes(x=id, y = rating)) + geom_point()
#ggplot(data = simpsons, aes(x=original_air_date, y = rating)) + geom_point()
```

- General negative trend as expected per the seasonal graph(s)
  - Start / end of season no notable patterns of yet granted it is somewhat hard to observe
  - i.e. season premiere vs finale (Presumably because of episodic nature and no overarching plotline)

#### Viewers

```{r Episode vs Viewer}
tempPlot <- ggplot(data=tempSimps, 
                   aes(x = id, y = viewers, col = as.factor(season), shape=as.factor(season))) + geom_point() + ylab("Viewer") + xlab("Episode Number") + ggtitle("Episode vs Viewer") + geom_line() + geom_line() + 
  scale_shape_manual(values = rep(0:14, 3)) + theme(legend.position = "none")

ggplotly(tempPlot)

#ggplot(data = simpsons, aes(x=id, y = rating)) + geom_point()
#ggplot(data = simpsons, aes(x=original_air_date, y = rating)) + geom_point()
```

- Can see more detailed info when zoomed in accordingly
  - However, graph is a little tight to get more consive view
- Overall trend does show that rating does plateau (as expected from seasonal boxplots)


- - - 

#### Season Specific

```{r Season Specific, echo=FALSE, include=FALSE}
# 
# get_season <- function(season_num) {
#   tempDF <- tempSimps %>% filter(season == season_num)
#   return(tempDF)
# }
# 
# s8 <- get_season(8)
# 
# set1 <- c()
# set2 <- c()
# set3 <- c()
# j
# plot_data_column = function (data) {
#     ggplot(data, aes_string(x = number_in_season, y = as.numeric(us_viewers_in_millions))) +
#         geom_point()
# }
#
# myplots <- lapply(1:12, plot_data_column, data = get_season(i))
# 
# for (i in 1:12) {
#   temp <- get_season(i)
#   tempGG <- ggplot(temp, aes(x = id, y = viewers)) + geom_point() +
#     ggtitle(glue('Season {i}'))
#   set1[[i]] = tempGG
# }
# 
# tempIndex = 1
# for (i in 13:24) {
#   temp <- get_season(i)
#   tempGG <- ggplot(temp, aes(x = id, y = viewers)) + geom_point() +
#     ggtitle(glue('Season {i}'))
#   set2[[tempIndex]] = tempGG
#   tempIndex = tempIndex + 1
# }
# 
# tempIndex = 1
# for (i in 24:max(simpsons$season)) {
#   temp <- get_season(i)
#   tempGG <- ggplot(temp, aes(x = id, y = viewers)) + geom_point() +
#     ggtitle(glue('Season {i}'))
#   set3[[tempIndex]] = tempGG
#   tempIndex = tempIndex + 1
# }
# 
# # grid.arrange(set1[[1]], set1[[2]], set1[[3]], set1[[4]], ncol=2, nrow=2)
# 
# 
# subplot(set1[[1]], set1[[2]])
# 
# ggplotly(grid.arrange(grobs = set1, ncol = 3, nrow=4))
# grid.arrange(grobs = set2, ncol = 3, nrow=4)
# grid.arrange(grobs = set3, ncol = 3, nrow=4)
# #
# for (i in 1:10) {
#   print(set1[i])
# }
# s1 <- get_season(1)
# ggplot(s1, aes(x = id, y = as.numeric(us_viewers_in_millions))) + geom_point()
# head(s1)
```













```{r (old) Viewers vs Episode, class.source = 'fold'}
# (oldViewers <- ggplot(data = old_simps, aes(x = id, y = us_viewers_in_millions, col = as.factor(season))) + geom_point() + ylab("Viewers") + xlab("Episode Number") + ggtitle("Num Viewers vs Episode") + theme(legend.position = "none"))
```

- ~~~Changing predictor variable to viewer count is interesting as pattern does not match way I thought it would~~~
  - ~~~i.e. Lower ratings = lower viewer, but need to reflect about how sample size would affect ratings~~~
  - Viewer count is in _millions_ 
  - ~~~Maybe an on/off season schedule? (Spring/Summer vs Fall/Winter)~~~
  - NOTE: Some data is missing are some `us_viewers_in_millions` values hard-coded as `N/A` (char class)
    - 159, 160, 172
- Old graph where data not cleaned, note the peculiar pattern

```{r Viewers Fixed}
simpsons$viewers <- as.numeric(simpsons$viewers)
(refactoredViewer <- ggplot(data = simpsons, aes(x = id, y = viewers, col = as.factor(season))) + 
  geom_point() + ylab("Viewers") + xlab("Episode Number") + ggtitle("Num Viewers vs Episode") +
  scale_y_continuous(name="Viewers (in mil)", limits=c(0,30)))


ggplot(data = tempSimps, aes(x = id, y = viewers, col = as.factor(season))) + geom_point()
```

- Upon refactoring viewer count to numeric trend does indeed follow expectations

```{r THOH}
#special_eps < simpsons %>% filter(simpsons, number_in_series = [4, 50, 20, 203])
THOH <- simpsons %>% filter(str_detect(title, 'Treehouse of Horror'))
ggplot(data = simpsons, aes(x = id, y = rating)) + geom_point() +
  geom_point(data = THOH, aes(x = id, y = rating, col = "")) + 
  ggtitle("Treehouse of Horror Specials")

ggplot(data = simpsons, aes(x = id, y = as.numeric(viewers))) + geom_point() +
  geom_point(data = THOH, aes(x = id, y = as.numeric(viewers), col = "")) + 
  ggtitle("Treehouse of Horror Specials")
```
- In general Treehouse of Horror (THOH) episodes seem to be on higher end of rating per season
- THOH 31 is lowest, and lower even within the immediate range / season




## Writers / Directors

```{r Writers}
length(unique(simpsons$writers))
length(unique(simpsons$directors))

mean(nchar(simpsons$writers))
avgWriter <- simpsons[which(nchar(simpsons$writers) > 16.6069 ),]

ggplot(simpsons, aes(writers, fill = writers)) + geom_bar() +
  coord_polar("y", start = 0)
```

- 187 unique writers records (Some entries have multiple people, i.e. outliers mentioned above - 2 part episodes)
  - The way data obtained caused errors/inconsistencies within data
    - Scrapped using different delimiters (&, spaces, or no space at all)
    - Some entries scraped including "Story by:" and/ok "Teleplay by:"
      - Different semantics based off the WGA (How did WGA Strike of 07 affect quality)
- 57 unique director listings 


https://stat.ethz.ch/R-manual/R-devel/library/base/html/nchar.html

## Data Analysis

## Predictive Model

## Sentiment Analysis off Descriptions



#----------------------



```{r asdf1, include=FALSE}
# tempSeason1 <- get_season(1)
# tempSeason2 <- get_season(2)
#
#
# ggplot(data=simpsons, aes(x = original_air_date, y = rating)) + geom_point(aes(col = as.factor(season))) + geom_smooth(data = simpsons, method=lm, se=FALSE, fullrange=FALSE, aes(group=as.factor(season), col="#000000")) + geom_smooth(data = get_season(1), aes(x=original_air_date, y = rating)) +
#   geom_smooth(data = get_season(2), aes(x=original_air_date, y = rating, col="2"))

```


# Misc Resources
- [strsplit()](https://stackoverflow.com/questions/8920145/count-the-number-of-all-words-in-a-string)
- [Collapsable Elements](https://stackoverflow.com/questions/52576626/rmarkdown-collapsible-panel)
- [paged_table()](https://stackoverflow.com/questions/23188900/display-print-all-rows-of-a-tibble-tbl-df)
